GVN-Hoist: Hoisting Computations from Branches

### 2016-08-07 hoist.tex
overview

\begin{verbatim}
float foo(float d, float min, float max, float a)
{
  float tmin;
  float tmax;

  float inv = 1.0f / d;
  if (inv >= 0) {
    tmin = (min - a) * inv;
    tmax = (max - a) * inv;
  } else {
    tmin = (max - a) * inv;
    tmax = (min - a) * inv;
  }
  return tmax + tmin;
}
\end{verbatim}

In this program the computations of tmax and tmin and identical in
both the branches but they cannot be evaluated before inv. Since inv
depends on a division operation which is generally more expensive than
the addition, subtraction and multiplication operations as required to
evaluate tmax and tmin. The total latency of computation across each
branch is:

@@

But if we hoist the computation of tmax and tmin outside the
conditionals, the C code version would look like this:
\begin{verbatim}
float foo(float d, float min, float max, float a)
{
  float tmin;
  float tmax;

  tmin' = (min - a);
  tmax' = (max - a);

  float inv = 1.0f / d;
  tmin' = tmin' * inv;
  tmax' = tmax' * inv;

  if (inv >= 0) {
    tmin = tmin';
    tmax = tmax';
  } else {
    tmin = tmax';
    tmax = tmin';
  }

@@

In this code the two subtractions and the div operations can execute
in parallel. So the total number of cycles will be max(O(div), O(sub)) + O(mul)
= O(div) + O(mul); since O(div) is usually greater than O(sub)

@@

\begin{verbatim}
float foo(float d, float min, float max, float a)
{
  float tmin;
  float tmax;

  tmin' = (min - a);
  tmax' = (max - a);

  float inv = 1.0f / d;
  tmin' = tmin' * inv;
  tmax' = tmax' * inv;

  return tmax' + tmin';
}
\end{verbatim}

>>>>

\begin{verbatim}
float foo(float d, float min, float max, float a)
{
  float tmin, tmax, inv;

  inv = 1.0f / d;
  if (inv >= 0) {
    tmin = (min - a) * inv;
    tmax = (max - a) * inv;
  } else {
    tmin = (max - a) * inv;
    tmax = (min - a) * inv;
  }
  return tmax + tmin;
}
\end{verbatim}

In this program the computations of tmax and tmin are identical to the
computations of tmin and tmax of sibling branch respectively. Both tmax and tmin
depends on inv which depends on a division operation which is generally more
expensive than the addition, subtraction and multiplication operations. The
total latency of computation across each branch is:

@@

Now if the computation of tmax and tmin are hoisted outside the
conditionals, the C code version would look like this:
\begin{verbatim}
float foo(float d, float min, float max, float a)
{
  float tmin, tmax, tmin1, tmax1, inv;

  tmin1 = (min - a);
  tmax1 = (max - a);

  inv = 1.0f / d;
  tmin1 = tmin1 * inv;
  tmax1 = tmax1 * inv;

  if (inv >= 0) {
    tmin = tmin1;
    tmax = tmax1;
  } else {
    tmin = tmax1;
    tmax = tmin1;
  }

@@

In this code the two subtractions and the division operations can be executed in
parallel because there are no dependencies among them. So the total number of
cycles will be max(O(div), O(sub)) + O(mul) = O(div) + O(mul); since O(div) is
usually much greater than O(sub) \cite{x86,aarch64}

@@

\begin{verbatim}
float foo(float d, float min, float max, float a)
{
  float tmin1, tmax1, inv;

  tmin1 = (min - a);
  tmax1 = (max - a);

  inv = 1.0f / d;
  tmin1 = tmin1 * inv;
  tmax1 = tmax1 * inv;

  return tmax1 + tmin1;
}
\end{verbatim}
